---
title: Using R dataset
description: |
  A simple example of exploring the mtcars dataset in R.
output: 
    html_document:
        self_contained: true
        toc: true
        toc_depth: 3
        toc_float:
            collapsed: false
            smooth_scroll: true
---

# Explore the `mtcars` dataset

## About the dataset

The `mtcars` data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (197374 models).

> Source: Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391411.

## Load the dataset

Load the `mtcars` dataset.

```{r}
# Load data
data(mtcars)
```

## Explore the dataset

Description of variables:

- mpg: Miles/(US) gallon
- cyl: Number of cylinders
- disp: Displacement (cu.in.)
- hp: Gross horsepower
- drat: Rear axle ratio
- wt: Weight (1000 lbs)
- qsec: 1/4 mile time
- vs: V/S
- am: Transmission (0 = automatic, 1 = manual)
- gear: Number of forward gears
- carb: Number of carburetors

```{r}
# Print object structure
str(mtcars)

# Print the first 10 rows
head(mtcars, 10)
```

Print observations and variables.

```{r}
# Number of observations
nrow(mtcars)

# Number of variables
ncol(mtcars)
```

Summaries the dataset.

```{r]
summary(mtcars)
```

## Create plots

### Histogram

Create a histogram of gross horsepower (hp).

```{r}
hist(mtcars$hp, xlab = "Gross horsepower", main = "Histogram of gross horsepower")
```

### Density plot

Create a density plot of weight.

```{r}
plot(density(mtcars$wt), xlab = "Weight (1000 lbs)", main = "Density plot of weight")
```

### Pie chart

Create a pie chart of the type of transmission.

```{r}
pie(table(mtcars$am), labels = c("automatic", "manual"), main = "Pie chart of transmission")
```

### Barplot

Create a barplot of number of forward gears.

```{r}
barplot(table(mtcars$gear), xlab = "Number of forward gears", 
	main = "Barplot of number of forward gears")
```

# Session information

```{r}
sessionInfo()
```

